---
title: 'R Programming: Week 2 - Functions'
output:
  word_document: default
  html_document:
    df_print: paged
---
#Objectives
Write an if-else expression
Write a for loop, a while loop, and a repeat loop
Define a function in R and specify its return value [see Functions Part 1 and Part 2]
Describe how R binds a value to a symbol via the search list
Define what lexical scoping is with respect to how the value of free variables are resolved in R
Describe the difference between lexical scoping and dynamic scoping rules. 
Convert a character string representing a date/time into an R datetime object. [see Dates and Times]


##Functions

Example 1: 
```{r}
add2 <- function(x , y ) {
   x + y
}
add2(3, 5)
```

Example 2:
```{r}
above10 <- function(x){ 
  use <- x > 10
  x[use]
}

above <- function(x,n) { 
    use <- x > n
x[use]
}
x <- 1:20
above(x, 12)
```

Example 3:
```{r}
columnmean <- function(y){
  nc <- ncol(y)
  means <- numeric(nc)
  for(i in 1:nc) {
    means[i] <- mean(y[,i], na.rm = removeNA)
  }
  means
}
```
<input dataset/> and colmeans(airquality, FALSE)

##Basics
Functions are sed to store R Objects
  -Considered to be first class objects
  -Include formal argument
 
 Functions arguments can be partially matched.
  - Order of operations 
    - Check for exact match for a name argument
    - Check for a partial match
    - Check for a positional match
    
    
Example 4: Defining a function
```{r}
f <- function(a, b = 1, c = 2, d = NULL) {
}
f
```

Example 5: Lazy Evaluation (arguments not being properly defined)
```{r}
f <- function(a ,b) {
   a^2
}
f(2)
```

Example 6: No Good Lazy Evaluation
``{r}
f <- function (a,b) {
  print(a)
  print(b)
}
f(45)
```

... Argument
- Indicates a variable number of arguments and is often used to not copy entire argument list of the original function
- Is necessary when the number of arguments passed to the function cannot be known in advance.

Example 7: ... Argument
```{r}
myplot <- function(x,y,type = "1",..) {
  plot(x,y,type=type,...)
}

```