---
title: "R Programming: Week 2 - Controlling Flows"
output: pdf_document
---
#Objectives
Write an if-else expression
Write a for loop, a while loop, and a repeat loop
Define a function in R and specify its return value [see Functions Part 1 and Part 2]
Describe how R binds a value to a symbol via the search list
Define what lexical scoping is with respect to how the value of free variables are resolved in R
Describe the difference between lexical scoping and dynamic scoping rules. 
Convert a character string representing a date/time into an R datetime object. [see Dates and Times]


##If/Else
If/Else tests conditions.

Basics

Example 1:
```{r}
x <- 4
if(x > 3) { y <- 10} else y <- 0;
y
```

Example 2:
```{r}
x=0
y <- if(x > 3) {10
} else {
 0
}
```

Example 1 and 2 produce the same output.
 
Can use just if/if statements


## For Loops
For loops execute a loop a fixed number of times.

Example 1: Basic
```{r}
for(i in 1:10) {
 print(i)
};
```

Example 2: All do the same thing
```{r}
x <- c("a", "b", "c", "d")
for(i in 1:4) {
 print(x[i])
}

for(i in seq_along(x)) {
 print(x[i])
}

for(letter in x) {
 print(letter)
}

for(i in 1:4) print(x[i]);
```

##While Loops
While Loops execute a loop while a condition is true

Example 1: Basic While
```{r}
count <- 0
while(count < 10) {
 print(count)
 count <- count + 1
};
```

Example 2: Multiple Conditions
```{r}
z <- 5
while(z >= 3 && z <= 10) { print(z) 
  coin <- rbinom(1, 1, 0.5)
 if(coin == 1) { ## random walk
 z <- z + 1
 } else {
 z <- z - 1
 }}
```

##Repeat, Next, and Break
-Repeat executes an infinite loop
  -Is not heavily used in R generally
-Next skips an iteration of a loop
-Break breaks the execution of a loop
  - Used to exit a function

Example 1: Next
```{r}
for(i in 1:100) {
 if(i <= 20) {
 ## Skip the first 20 iterations
 next
 }
 ## Do something here
};
```

Helpful hints
-Avoid infinite loops
-If working in commane line, use *apply functions.


